\Title

### 2021-09-10 latex/nserc/introduction.tex
Modified the main text based on OGS_2021, the paper, and my thesis.

Now imagine that the kiiker is a robot, and their creator is teaching them to
swing like a human.
If the roboticist had studied classical control theory, they would plot 
a kiiker's squat depth as a trajectory over time and tell the robot to
synchronize with this trajectory.
In an ideal world, this technique would work perfectly; unfortunately, it is
also entirely unnatural. 
After all, humans on swings do not have an internal stopwatch telling them when
to squat.
Rather, they adjust how much they bend their knees based on the current angle of
the swing along with their direction of motion.
This position-velocity adjustment allows humans to correct for external
disturbances (like strong winds or enthusiastic swing pushers). 
Figure \ref{fig:swing-pos-vel} shows this adjustment process:
a person will squat when they reach the peak of their swing, and
stand when they reach the fastest point at the bottom
\cite{pumping_swing_standing_squatting}.
Even if the robot could perfectly track a time-based trajectory, an external
disturbance may affect where squatting and standing occurs to the point that
the swing actually slows down.

@@

The modern concept of virtual nonholonomic constraints was described by
\citet{nhvc_dynamic_walking} in 2015, though there are references to more
primitive versions going back as early as the the year 2000
\cite{vnhc_human_robot_cooperation}.
Virtual nonholonomic constraints have been of most notable use in
bipedal locomotion 
\cite{nhvc_gait_optimization,output_nhvc_bipedal_control},
where they have shown marked improvements in the robustness of walking gaits
when compared to previous control techniques.
They have also been used for error-reduction in time-delayed teleoperation
\cite{vnhc_time_delay_teleop} and in the field of human-robot interaction
\cite{psd_based_vnhc_redundant_manipulator,haptic_vnhc}.
\citet{hybrid_zero_dynamics_bipedal_nhvcs} have derived the 
zero dynamics for hybrid mechanical systems under virtual nonholonomic
constraints.
In particular, they study a class of constraints formed by taking an affine
sum of holonomic functions with nonholonomic B\'{e}zier polynomials.
They widen this class of constraints in \cite{nhvc_incline_walking} to improve
bipedal locomotion on variable-slope terrain, and generate constraints through
a specialized optimization routine.

>>>>

Now imagine that the kiiker is a robot, and their creator is teaching them to
swing like a human.
If the roboticist had studied classical control theory, they would plot 
a kiiker's squat depth as a trajectory over time and tell the robot to
synchronize with this trajectory.
In an ideal world, this technique would work perfectly; unfortunately, it is
also entirely unnatural. 
After all, humans on swings do not have an internal stopwatch telling them when
to squat.
Rather, they adjust how much they bend their knees based on the current angle of
the swing along with their direction of motion.
This position-velocity adjustment allows humans to correct for external
disturbances (like strong winds or enthusiastic swing pushers). 
Figure \ref{fig:swing-pos-vel} shows this adjustment process:
a person will squat when they reach the peak of their swing, and
stand when they reach the fastest point at the bottom
\cite{pumping_swing_standing_squatting}.
Even if the robot could perfectly track a time-based trajectory, an external
disturbance may affect where squatting and standing occur to the point that
the swing actually slows down.

@@

The modern concept of virtual nonholonomic constraints was described by
\citet{nhvc_dynamic_walking} in 2015, though there are references to more
primitive versions going back as early as the year 2000
\cite{vnhc_human_robot_cooperation}.
Virtual nonholonomic constraints have been of most notable use in
bipedal locomotion 
\cite{nhvc_gait_optimization,output_nhvc_bipedal_control},
where they have shown marked improvements in the robustness of walking gaits
when compared to previous control techniques.
They have also been used for error-reduction in time-delayed teleoperation
\cite{vnhc_time_delay_teleop} and in the field of human-robot interaction
\cite{psd_based_vnhc_redundant_manipulator,haptic_vnhc}.
\citet{hybrid_zero_dynamics_bipedal_nhvcs} have derived the 
zero dynamics for hybrid mechanical systems under virtual nonholonomic
constraints.
In particular, they study a class of constraints formed by taking an affine
sum of holonomic functions with nonholonomic B\'{e}zier polynomials.
They widen this class of constraints in \cite{nhvc_incline_walking} to improve
bipedal locomotion on variable-slope terrain, and generate constraints through
a specialized optimization routine.

### 2021-09-10 latex/nserc/main.tex
Deleting extraneous data and just writing the nserc proposal
Modified the main text based on OGS_2021, the paper, and my thesis.

@@

%---------- Title -----------%
\title{Ontario Graduate Scholarship: Plan of Study}
\author{Adan Moran-MacDonald}
\date{}
\maketitle

@@

When a gymnast on a horizontal bar wants to swing up to do backflips, they
typically perform a technique called a ``giant".
This technique involves the gymnast swinging their legs based on their current
speed and position, thereby generating enough momentum to rotate around the bar
\cite{pendulum_length_giant_gymnastics}.

Standard practice in robotics begins with finding a trajectory over time
replicating a desired behaviour, then using the robot's actuators to track this
time-based trajectory. 
While this often works for simple systems, it can cause timing delays, it is not
robust to external disturbances, and it is not what humans do.
A more natural approach is to constrain the robot's actuators according to
some function of speed and position.
This type of behaviour is well described by the method of virtual constraints.

There are two types of constraints on a mechanical system:
\textit{holonomic} constraints restrict position (e.g. a snake robot can slither
but not fly),
while \textit{nonholonomic} constraints restrict position and
velocity (e.g. cars cannot slide sideways).
One can design a specific constraint for a robot and enforce it using the
actuators.
Since this constraint is not enforced by physics, it is known as a ``virtual"
constraint
\cite{vhcs_for_el_systems}.
Virtual holonomic constraints (VHCs) have been used to control walking robots
\cite{stable-walking}, autonomous bicycles \cite{bicycle},
gymnastics robots \cite{xingbo-thesis}, and snake robots \cite{snake-robot}.
Virtual nonholonomic constraints (VNHCs) have been used for human-robot
cooperation \cite{vnhc-human-robot-coop} and for improving robustness of walking
robots \cite{vnhc-biped-robot,hybrid_zero_dynamics_bipedal_nhvcs}, among other
applications.

@@

Virtual constraints have proven useful for generating biologically-inspired
behaviour in robotic systems.
In my master's thesis, I demonstrated that VNHCs can also inject energy into
robots:
I designed a VNHC which enabled a gymnastics robot to gain energy
and perform backflips on a horizontal bar \cite{my-thesis}.
Using VNHCs for \textit{energy injection} is beneficial because it allows robots to
safely increase their momentum.
Likewise, designing VNHCs for \textit{energy dissipation} allows robots to
safely slow down in a realistic and reliable manner.

We propose to study the energy injection and dissipation properties of VNHCs.
First, we will find mathematical conditions under which VNHCs are guaranteed to
inject/dissipate energy in abstract mechanical systems.
Then we will design a means of safely transitioning between two different VNHCs.
Finally, we will test the theory on a gymnastics robot by designing a gymnastics
routine which is performed solely by transitioning between different virtual
constraints.

>>>>

%---------- Add name to Header ----------%
\usepackage{atbegshi,picture}
\AtBeginShipout{\AtBeginShipoutUpperLeft{%
        \put(\dimexpr\paperwidth-5.5cm\relax,-1.2cm){\makebox{\setlength{\fboxrule}{0pt}\framebox{Adan
        Moran-MacDonald}}}%
}}

@@

%---------- Title -----------%
\title{Energy Regulation for Biologically-inspired Robotics}
\author{}
\date{}
\maketitle

\vspace{-1cm} % Move the following text up
\textbf{What is your area of research? What sort of problems do you want to
    address? WHy is the area important, and why are the applications important?
    What is special about your approach, and how is it different from
    existing/competing approaches? What is a sample application of the outcomes
of your thesis?}

@@

When a gymnast wants to do backflips around a horizontal bar, they start by
performing a technique called a ``giant" \cite{usagym_giant}. 
The gymnast swings their legs based on their current speed and position, thereby
generating enough momentum to rotate around the bar.

Now imagine the gymnast is a robot, and their creator is teaching them to act
like a human.
If the roboticist had studied classical control theory, they would plot a
human gymnast's hip angle as a trajectory over time and tell the robot to synchronize
its legs with this trajectory.
While this approach often works for simple systems, it is not great for
biologically-inspired robots: it can cause timing delays, it is not robust to
many external disturbances, and it is \textit{not what humans do}.
After all, human gymnasts do not have an internal stopwatch telling them when to
extend or retract their legs.
Instead, existing research suggests that human gymnasts actually move their legs
as a function of their body angle and velocity
\cite{pendulum_length_giant_gymnastics}.
Rather than using time-based motion, a clever roboticist would attempt to
emulate this more natural human behaviour.
% by driveing robot's legs according to some function of its torso speed and position.
This type of behaviour is well described by the method of virtual constraints.

There are two types of constraints on a mechanical system:
\textit{holonomic} constraints restrict position (e.g. a snake robot can slither
but not fly),
while \textit{nonholonomic} constraints restrict position and
velocity (e.g. cars cannot slide sideways).
One can design a constraint they want their robot to satisfy and enforce it
using the robot's actuators.
Since this constraint is not enforced by physics, it is known as a ``virtual"
constraint \cite{vhcs_for_el_systems}.
Virtual holonomic constraints (VHCs) have been used to control walking robots
\cite{stable-walking}, autonomous bicycles \cite{bicycle},
gymnastics robots \cite{xingbo-thesis}, and snake robots \cite{snake-robot}.
Virtual nonholonomic constraints (VNHCs) have been used for human-robot
cooperation \cite{vnhc-human-robot-coop} and for improving robustness of walking
robots \cite{vnhc-biped-robot,hybrid_zero_dynamics_bipedal_nhvcs}, among other
applications.

@@

Virtual constraints have proven useful for generating biologically-inspired
behaviour in robotic systems.
In my master's thesis, I demonstrated that VNHCs can inject/dissipate energy
into/from robots:
I designed a VNHC which enabled a gymnastics robot to gain energy
and perform backflips on a horizontal bar \cite{my-thesis}.
Using VNHCs for \textit{energy injection} is beneficial because it allows robots to
safely increase their momentum.
Likewise, designing VNHCs for \textit{energy dissipation} allows robots to
safely slow down in a realistic and reliable manner.

We propose to study the energy injection and dissipation properties of VNHCs.
First, we will find mathematical conditions under which VNHCs are guaranteed to
inject/dissipate energy in arbitrary mechanical systems.
Then we will design a means of transitioning between two different VNHCs,
enabling us to regulate energy in a provably safe manner.
Finally, we will test the theory on two robots: a gymnastics robot and a robotic
monkey.
For the gymnastics robot (which is already built), we will design a gymnastics
routine which is performed solely by transitioning between different virtual
constraints.
After this, we will build the robotic monkey and design virtual constraints
which allow it to swing on monkey bars, just like real monkeys swing through
trees.

### 2021-09-10 latex/nserc/paper.tex
Modified the main text based on OGS_2021, the paper, and my thesis.

%/========== Problem Formulation ==========/%
\section{Problem Formulation}\label{sec:problem-formulation}
We will use the simplified acrobot model in Figure
\ref{fig:acrobot}, where we assume the torso and leg rods are of
equal length \(l\) with equal point masses \(m\) at the tips.
The acrobot's configuration is described in generalized coordinates
\((q_u,q_a)\) on the configuration manifold 
\(\mathcal{Q} = \Sone \times \Sone\), where \(q_u\) is unactuated and 
\(q_a\) is actuated.
A real gymnast cannot swing their legs in full circles, though they
are usually flexible enough to raise them parallel to the floor;
hence, we assume the leg angle \(q_a\) lies in \([-Q_a, Q_a]\) for some
\(Q_a \in [\frac{\pi}{2}, \pi[\). 
We also ignore any dissipative forces.

The acrobot has inertia matrix 
\(M\), potential function \(V\) (with respect to
the horizontal bar), and input matrix \(B\) given as follows:
\begin{align}\label{eqn:acrobot-inertia}
    M(q) &= \begin{bmatrix}
        ml^2\left(3+2\cos(q_a)\right) & 
        ml^2\left(1+\cos(q_a)\right) \\
        ml^2\left(1+\cos(q_a)\right) &
        ml^2
    \end{bmatrix} 
    , \\
    \label{eqn:acrobot-potential}
    V(q) &= -mgl\left(2\cos(q_u)+\cos(q_u+q_a)\right)
    , \\
    \label{eqn:acrobot-B}
    B &= [0;1]
    .
\end{align}
For reasons that will become clear in later sections of this article, we will
use Hamiltonian mechanics to derive the dynamics of the acrobot. 
For this we require the conjugate of momenta, \(p = (p_u,p_a) = M(q)\dot{q}\).
The dynamics of the acrobot in \((q,p)\) coordinates are given in
\eqref{eqn:acrobot-hamiltonian}.
For shorthand, we write \(c_u := \cos(q_u)\), \(c_a := \cos(q_a)\), and 
\(c_{ua} := \cos(q_u + q_a)\); likewise, \(s_u := \sin(q_u)\), 
\(s_a := \sin(q_a)\), and \(s_{ua} := \sin(q_u + q_a)\).
\begin{align}\label{eqn:acrobot-hamiltonian}
    \mathcal{H}(q,p) &= \frac{1}{2}p\tpose \Minv(q) p -
    mgl\left(2 c_u + c_{ua}\right)
    , \\
     &\begin{cases}
        \dot{q} = \Minv(q) p 
        ,\\
        \dot{p}_u = -mgl\left(2s_u + s_{ua}\right) 
        ,\\
        \dot{p}_a =-\frac{1}{2}p\tpose \nabla_{q_a}\Minv(q) p
        - mgl s_{ua} + \tau.
    \end{cases} \nonumber
\end{align}
The control input is a force \(\tau \in \R\) affecting only the dynamics of
\(p_a\), representing a torque acting on the hip joint.
Let us now define what it means for a mechanical system to gain or lose energy.

@@

The following theorem requires that \(\mathcal{Q} = \R^n\) so that the
transformation \(\tilde{q} = \mathbf{B}q\) is well defined.

@@

Now imagine the person's torso is affixed to the swing's rope so they are
always upright. 
Imagine further that the swing has no seat at all, allowing the person to extend
their legs beneath them. 
This position is identical to that of a gymnast on a bar.

The acrobot's legs are rigid rods which cannot retract, so we emulate the person
on a swing by pivoting the legs toward the direction of motion. 
Since a person lifts their legs higher at faster speeds, the acrobot's legs should
pivot to an angle proportional to the swing's speed.
Because the direction of motion is entirely determined by \(p_u\), 
one VNHC which emulates this process is \(q_a = \bar{q}_a\arctan( I p_u)\),
displayed in Figure \ref{fig:qa-arctan}.
Here, \(\bar{q}_a \in \, ]0,2 Q_a/\pi]\) and \(I \in \R\) are 
design parameters.
This constraint does not perfectly recreate giant motion, during which
the gymnast's legs are almost completely extended \cite{usagym_giant}.
It instead pivots the legs partially during rotations.
However, the behaviour is similar enough that this constraint will
still inject energy into the acrobot.
Our final VNHC is
\begin{equation}\label{eqn:acrobot-constraint}
    h(q,p) = q_a - \bar{q}_a \arctan(I p_u)
    .
\end{equation}

@@

\textbf{Rotation Regulation}: one chooses a desired rotation rate 
\(p_\text{des} > 0\) and a hysteresis value \(\delta \in [0,1]\).
Each time the orbit crosses the \(p_u\)-axis (i.e. when \(q_u = 0\)), the
supervisor changes which VNHC is enforced as follows:
\begin{itemize}
    \item If \(|p_u| < (1-\delta)p_\text{des}\), enable the injection VNHC by
        setting \(I > 0\) in \eqref{eqn:acrobot-constraint}.
    \item If \(|p_u| > (1+\delta)p_\text{des}\), enable the dissipation VNHC by
        replacing \(I > 0\) with \(-I < 0\) in \eqref{eqn:acrobot-constraint}.
    \item If \((1-\delta)p_\text{des} \leq |p_u| \leq (1+\delta)p_\text{des}\),
        extend the leg fully by setting \(q_a = 0\).
        For these simulations, we assume this can be done instantaneously,
        though in practice this is impossible.
\end{itemize}
Note that all orbits of the acrobot must cross the \(p_u\) axis, 
so this method will stabilize a rotation even if the acrobot is initialized in
\(\mathcal{O}_1\).
Also, the above supervisor is designed to regulate a rotation rate, but it
cannot regulate the rotation direction.

@@

We can extend this idea to find the nominal pendulum associated with our
physical acrobot, whose parameters are provided in Table
\ref{tab:acrobot-parameters}.
Setting \(I = 0\) in \eqref{eqn:acrobot-constraint} for this acrobot
yields the following mechanical energy of the nominal pendulum:
\[
    E(q_u,p_u) \approx 396.5501 p_u^2 + 0.5997(1 - \cos(q_u))
    .
\]
Despite the change in model, the level set \(E_\pi\) with energy 
\(E(\pi,0)\) remains the boundary between oscillations and rotations
for this nominal pendulum.

@@

Finally, The Dynamixel servo motor does not have a torque control mode; instead,
we can assign the servo setpoint at iteration \(k \in \mathbb{Z}_{> 0}\)
via \(q_a^{k} = \arctan(I p_u^{k-1})\).
This negatively affects the stabilization to the constraint
manifold because we are introducing timing errors from the servo's built-in PID
controller.

@@

\begin{lemma}\label{lemma:energy-regulation-D}
    Consider the simplified acrobot \eqref{eqn:acrobot-hamiltonian} constrained
    by the VNHC \eqref{eqn:acrobot-constraint}, whose constrained dynamics are
    \eqref{eqn:acrobot-constrained-dynamics}.
    Let \(D \subset \SxR\) be open and 
    let \(]\underline{r}, \overline{r}[ \, \subset \R_{> 0}\).
    Suppose there exists a diffeomorphism 
    \begin{align*}
        T : D &\to ]\underline{r},\overline{r}[ \, \times \Sone
         \\
        (q_u,p_u) &\mapsto (r,\theta)
        ,
    \end{align*}
    yielding the new constrained dynamics \(\dot{r} = f_r(\theta,r,I)\),
    \(\dot{\theta} = f_\theta(\theta,r,I)\).
    Suppose that \(f_r(\theta,r,0) \equiv 0\) and \(f_\theta(\theta,r,0) > 0\).
    Let \(g(\theta,r,I) = f_r(\theta,r,I)/f_\theta(\theta,r,I)\), and assume that 
    \((\partial g/\partial r)(\theta,r,0) \equiv 0\).
    Assume also that \(b(\theta,r) := (\partial g/\partial I)(\theta,r,0)\)
    is \(2\pi\)-periodic in \(\theta\),
    and that there exists \(\epsilon > 0\) such that 
    \(S(r_0) := \int_0^{2\pi} b(\sigma,r_0)d\sigma \geq \epsilon\)
%    \[
%        S(r_0) := \int \limits_0^{2\pi} b(\sigma,r_0)d\sigma \geq \epsilon
%        ,
%    \]
    for all \(r_0 \in \, ]\underline{r},\overline{r}[\).
    Then there exists \(I^\star > 0\) small enough that the constrained dynamics
    gain energy on \(D\) for all \(I \in \,]0, I^\star]\), and lose energy on \(D\)
    for all \(I \in [-I^\star, 0[\).
\end{lemma}
\begin{proof}
    Since \(f_\theta(\theta,r,0) > 0\) and \(f_\theta\)
    is continuous in \(I\), there exists 
    \(I_1 > 0\) small enough that \(\dot{\theta} > 0\) for all
    \(I \in \, [-I_1,I_1]\).
    Hence, we can perform a time reparameterization \(t = t(\theta)\) to get a
    time-scaled pseudo-radius \(\hat{r}(\theta) := r(t(\theta))\).
    This reduces the system \((\dot{r},\dot{\theta})\) into the scalar
    time-varying ODE
    \begin{equation}\label{eqn:rhat-ode}
        \begin{cases}
            \hat{r}^\prime= \left.\frac{dr}{dt} \frac{dt}{d\theta}\right|_{r =
                \hat{r}} = 
            g(\theta,\hat{r},I)
            , \\
            \hat{r}(0) = r_0
            ,
        \end{cases}
    \end{equation}
    with state \(\hat{r}\), where now
    \(\theta\) represents time,
    prime denotes differentiation by \(\theta\), and
    \(r_0 \in \, ]\underline{r},\overline{r}[\) is the initial condition.

    Let \(\hat{r}(\theta,r_0,I)\) be the solution to \eqref{eqn:rhat-ode}.
    Taking the Taylor expansion of this solution around \(I = 0\) we get
    \(\hat{r}(\theta,r_0,I) = \hat{r}(\theta,r_0,0) + I\hat{r}_1(\theta,r_0) 
    + R(\theta,r_0,I)\). 
    Here, \(\hat{r}(\theta,r_0,0) \equiv r_0\) because
    \(g(\theta,\hat{r},0) \equiv 0\),
    the function \(\hat{r}_1(\theta,r_0)\) is the solution to
    \begin{equation}\label{eqn:r1-hat-prime}
        \begin{cases}
        \hat{r}_1^\prime = 
        \frac{\partial g}{\partial \hat{r}}(\theta,r_0,0)\hat{r}_1 +
        b(\theta,r_0)
        %\frac{\partial g}{\partial I}(\theta,r_0,0)
        , \\
        \hat{r}_1(0) = 0
        ,
    \end{cases}
    \end{equation}
    and the remainder term \(R(\theta,r_0,I)\) is \(O(I^2)\).

@@

Thankfully, we can still use Lemma \ref{lemma:energy-regulation-D} on an
appropriate subset of \(\mathcal{O}_1\).
To find this subset, first recall the discrete-time pseudo-radius dynamics
\eqref{eqn:r-discrete} from the proof of Lemma \ref{lemma:energy-regulation-D}.
Linearizing \eqref{eqn:r-discrete} at \(r_0 = 0\) yields
\[
    P^\prime(0) = 1 + I K Q^\prime(0) + R^\prime(2\pi,0,I)
    ,
\]
where the prime now denotes differentiation with respect to \(r\).
One can show that
\begin{align*}
    Q^\prime(0) = \int \limits_0^\pi 
    \lim \limits_{r_0 \to 0^+}\pdiff{a}{r}(\sigma,r_0) d\sigma
    = \frac{\pi}{2\sqrt{2}} > 0
    .
\end{align*}
Since \(R(2\pi,0,I)\) is \(O(I^2)\), so is
\(R^\prime(2\pi,0,I)\).
Hence, it can be written in the form
\(R^\prime(2\pi,0,I) = I^2 \tilde{R}(I)\) where \(\tilde{R}(I)\) is smooth and
zero at \(I = 0\).
Thus, there exists \(I_0 > 0\) such that 
\(I K Q^\prime(0) + I^2 \tilde{R}(I) > 0\) for all \(I \in \, ]0,I_0]\),
which implies \(P^\prime(0) > 1\) for all \(I \in \, ]0,I_0]\).
This means the origin is a repeller, so there exists some
(unknown) \(\mu > 0\) where \(]0,\mu]\) is negatively
invariant for \eqref{eqn:r-discrete}, and all solutions in this
interval flow through \(r = \mu\).
If \(I < 0\), we have \(P^\prime(0) < 1\) and the origin is an
asymptotically stable equilibrium for \eqref{eqn:r-discrete}.

>>>>

%/========== Problem Formulation ==========/%
\section{Problem Formulation}\label{sec:problem-formulation}
We will use the simplified acrobot model in Figure
\ref{fig:acrobot}, where we assume the torso and leg rods are of
equal length \(l\) with equal point masses \(m\) at the tips.
The acrobot's configuration is described in generalized coordinates
\((q_u,q_a)\) on the configuration manifold 
\(\mathcal{Q} = \Sone \times \Sone\), where \(q_u\) is unactuated and 
\(q_a\) is actuated.
We ignore dissipative forces in this model.

The acrobot has inertia matrix 
\(M\), potential function \(V\) (with respect to
the horizontal bar), and input matrix \(B\) given as follows:
\begin{align}\label{eqn:acrobot-inertia}
    M(q) &= \begin{bmatrix}
        ml^2\left(3+2\cos(q_a)\right) & 
        ml^2\left(1+\cos(q_a)\right) \\
        ml^2\left(1+\cos(q_a)\right) &
        ml^2
    \end{bmatrix} 
    , \\
    \label{eqn:acrobot-potential}
    V(q) &= -mgl\left(2\cos(q_u)+\cos(q_u+q_a)\right)
    , \\
    \label{eqn:acrobot-B}
    B &= [0;1]
    .
\end{align}
For reasons that will become clear later in this article, we
use Hamiltonian mechanics to derive the dynamics of the acrobot. 
For this we require the conjugate of momenta, \(p = (p_u,p_a) = M(q)\dot{q}\).
The dynamics of the acrobot in \((q,p)\) coordinates are given in
\eqref{eqn:acrobot-hamiltonian}.
For shorthand, we write \(c_u := \cos(q_u)\), \(c_a := \cos(q_a)\), and 
\(c_{ua} := \cos(q_u + q_a)\); likewise, \(s_u := \sin(q_u)\), 
\(s_a := \sin(q_a)\), and \(s_{ua} := \sin(q_u + q_a)\).
\begin{align}\label{eqn:acrobot-hamiltonian}
    \mathcal{H}(q,p) &= \frac{1}{2}p\tpose \Minv(q) p -
    mgl\left(2 c_u + c_{ua}\right)
    , \\
     &\begin{cases}
        \dot{q} = \Minv(q) p 
        ,\\
        \dot{p}_u = -mgl\left(2s_u + s_{ua}\right) 
        ,\\
        \dot{p}_a =-\frac{1}{2}p\tpose \nabla_{q_a}\Minv(q) p
        - mgl s_{ua} + \tau.
    \end{cases} \nonumber
\end{align}
The control input is a force \(\tau \in \R\) affecting only the dynamics of
\(p_a\), representing a torque acting on the hip joint.
Let us now define what it means for a mechanical system to gain or lose energy.

@@

The following theorem requires that \(\mathcal{Q} = \R^n\) so that the
coordinate transformation is well defined.

@@

Now imagine the person's torso is affixed to the swing's rope so they are
always upright. 
Imagine further that the swing has no seat, allowing the person to extend
their legs beneath them. 
This position is identical to that of a gymnast on a bar.

The acrobot's legs are rigid rods which cannot retract, so we emulate the person
on a swing by pivoting the legs toward the direction of motion. 
Since a person lifts their legs higher at faster speeds, the acrobot's legs should
pivot to an angle proportional to the swing's speed.
A real gymnast cannot swing their legs in full circles, though they
are usually flexible enough to raise them parallel to the floor;
hence, the VNHC must restrict the leg angle \(q_a\) 
to lie in \([-Q_a, Q_a]\) for some \(Q_a \in [\frac{\pi}{2}, \pi[\). 
Because the direction of motion is entirely determined by \(p_u\), 
one VNHC which emulates this process is \(q_a = \bar{q}_a\arctan( I p_u)\),
displayed in Figure \ref{fig:qa-arctan}.
Here, \(\bar{q}_a \in \, ]0,2 Q_a/\pi]\) and \(I \in \R\) are 
design parameters.

This constraint does not perfectly recreate giant motion, during which
the gymnast's legs are almost completely extended \cite{usagym_giant}.
It instead pivots the legs partially during rotations.
However, the behaviour is similar enough that this constraint will
still inject energy into the acrobot.
Our final VNHC is
\begin{equation}\label{eqn:acrobot-constraint}
    h(q,p) = q_a - \bar{q}_a \arctan(I p_u)
    .
\end{equation}

@@

\textbf{Rotation Regulation}: one chooses a desired rotation rate 
\(p_\text{des} > 0\) and a hysteresis value \(\delta \in [0,1]\).
Each time the orbit crosses the \(p_u\)-axis (i.e. when \(q_u = 0\)), the
supervisor changes which VNHC is enforced as follows:
\begin{itemize}
    \item If \(|p_u| < (1-\delta)p_\text{des}\), enable the injection VNHC by
        setting \(I > 0\) in \eqref{eqn:acrobot-constraint}.
    \item If \(|p_u| > (1+\delta)p_\text{des}\), enable the dissipation VNHC by
        replacing \(I > 0\) with \(-I < 0\) in \eqref{eqn:acrobot-constraint}.
    \item If \((1-\delta)p_\text{des} \leq |p_u| \leq (1+\delta)p_\text{des}\),
        extend the leg fully by setting \(q_a = 0\).
        In simulation we assume this can be done instantaneously,
        though in practice this takes time.
\end{itemize}
All orbits of the acrobot must cross the \(p_u\) axis, 
so this method will stabilize a rotation even if the acrobot is initialized in
\(\mathcal{O}_1\).
While the above supervisor is designed to regulate a rotation rate, it
cannot regulate the rotation direction.

@@

We can extend this idea to find the nominal pendulum associated with our
physical acrobot, whose parameters are provided in Table
\ref{tab:acrobot-parameters}.
Setting \(I = 0\) in \eqref{eqn:acrobot-constraint} for this acrobot
yields the mechanical energy of the nominal pendulum:
\[
    E(q_u,p_u) \approx 396.5501 p_u^2 + 0.5997(1 - \cos(q_u))
    .
\]
Despite the change in model, the level set \(E_\pi\) with energy 
\(E(\pi,0)\) remains the boundary between oscillations and rotations
for this nominal pendulum.

@@

Finally, the Dynamixel servo motor does not have a torque control mode; instead,
we can assign the servo setpoint at iteration \(k \in \mathbb{Z}_{> 0}\)
via \(q_a^{k} = \arctan(I p_u^{k-1})\).
This negatively affects the stabilization to the constraint
manifold because we are introducing timing errors from the servo's built-in PID
controller.

@@

\begin{lemma}\label{lemma:energy-regulation-D}
    Consider the simplified acrobot \eqref{eqn:acrobot-hamiltonian} constrained
    by the VNHC \eqref{eqn:acrobot-constraint}, whose constrained dynamics are
    \eqref{eqn:acrobot-constrained-dynamics}.
    Let \(D \subset \SxR\) be open and 
    let \(]\underline{r}, \overline{r}[ \, \subset \R_{> 0}\).
    Suppose there exists a diffeomorphism 
    \begin{align*}
        T : D &\to ]\underline{r},\overline{r}[ \, \times \Sone
         \\
        (q_u,p_u) &\mapsto (r,\theta)
        ,
    \end{align*}
    yielding the new constrained dynamics \(\dot{r} = f_r(\theta,r,I)\),
    \(\dot{\theta} = f_\theta(\theta,r,I)\).
    Suppose that \(f_r(\theta,r,0) \equiv 0\) and \(f_\theta(\theta,r,0) > 0\).
    Let \(g(\theta,r,I) := f_r(\theta,r,I)/f_\theta(\theta,r,I)\), and assume that 
    \((\partial g/\partial r)(\theta,r,0) \equiv 0\).
    Assume also that \(b(\theta,r) := (\partial g/\partial I)(\theta,r,0)\)
    is \(2\pi\)-periodic in \(\theta\),
    and that there exists \(\epsilon > 0\) such that 
    \(S(r_0) := \int_0^{2\pi} b(\sigma,r_0)d\sigma \geq \epsilon\)
%    \[
%        S(r_0) := \int \limits_0^{2\pi} b(\sigma,r_0)d\sigma \geq \epsilon
%        ,
%    \]
    for all \(r_0 \in \, ]\underline{r},\overline{r}[\).
    Then there exists \(I^\star > 0\) small enough that the constrained dynamics
    gain energy on \(D\) for all \(I \in \,]0, I^\star]\), and lose energy on \(D\)
    for all \(I \in [-I^\star, 0[\).
\end{lemma}
\begin{proof}
    Since \(f_\theta(\theta,r,0) > 0\) and \(f_\theta\)
    is continuous in \(I\), there exists 
    \(I_1 > 0\) small enough that \(\dot{\theta} > 0\) for all
    \(I \in \, [-I_1,I_1]\).
    Hence, we can perform a time reparameterization \(t = t(\theta)\) to get a
    time-scaled pseudo-radius \(\hat{r}(\theta) := r(t(\theta))\).
    This reduces the system \((\dot{r},\dot{\theta})\) into the scalar
    time-varying ODE
    \begin{equation}\label{eqn:rhat-ode}
        \begin{cases}
            \hat{r}^\prime= \left.\frac{dr}{dt} \frac{dt}{d\theta}\right|_{r =
                \hat{r}} = 
            g(\theta,\hat{r},I)
            , \\
            \hat{r}(0) = r_0
            ,
        \end{cases}
    \end{equation}
    with state \(\hat{r}\), where now
    \(\theta\) represents time,
    prime denotes differentiation by \(\theta\), and
    \(r_0 \in \, ]\underline{r},\overline{r}[\) is the initial condition.

    Let \(\hat{r}(\theta,r_0,I)\) be the solution to \eqref{eqn:rhat-ode}.
    Taking the Taylor expansion of this solution around \(I = 0\) yields
    \(\hat{r}(\theta,r_0,I) = \hat{r}(\theta,r_0,0) + I\hat{r}_1(\theta,r_0) 
    + R(\theta,r_0,I)\). 
    Here, \(\hat{r}(\theta,r_0,0) \equiv r_0\) because
    \(g(\theta,\hat{r},0) \equiv 0\),
    the function \(\hat{r}_1(\theta,r_0)\) is the solution to the linear
    time-varying ODE
    \begin{equation}\label{eqn:r1-hat-prime}
        \begin{cases}
        \hat{r}_1^\prime = 
        \frac{\partial g}{\partial \hat{r}}(\theta,r_0,0)\hat{r}_1 +
        b(\theta,r_0)
        %\frac{\partial g}{\partial I}(\theta,r_0,0)
        , \\
        \hat{r}_1(0) = 0
        ,
    \end{cases}
    \end{equation}
    and the remainder term \(R(\theta,r_0,I)\) is \(O(I^2)\).

@@

Thankfully, we can still use Lemma \ref{lemma:energy-regulation-D} on an
appropriate subset of \(\mathcal{O}_1\).
To find this subset, first recall the discrete-time pseudo-radius dynamics
\eqref{eqn:r-discrete} from the proof of Lemma \ref{lemma:energy-regulation-D}.
Linearizing \eqref{eqn:r-discrete} at \(r_0 = 0\) yields
\[
    P^\prime(0) = 1 + I K Q^\prime(0) + R^\prime(2\pi,0,I)
    ,
\]
where the prime now denotes differentiation with respect to \(r\).
One can show that
\begin{align*}
    Q^\prime(0) = \int \limits_0^\pi 
    \lim \limits_{r_0 \to 0^+}\pdiff{a}{r}(\sigma,r_0) d\sigma
    = \frac{\pi}{2\sqrt{2}} > 0
    .
\end{align*}
Since \(R(2\pi,0,I)\) is \(O(I^2)\), so is
\(R^\prime(2\pi,0,I)\).
Hence, it can be written in the form
\(R^\prime(2\pi,0,I) = I^2 \tilde{R}(I)\) where \(\tilde{R}(I)\) is smooth,
bounded, and zero at \(I = 0\).
Thus, there exists \(I_0 > 0\) such that 
\(I K Q^\prime(0) + I^2 \tilde{R}(I) > 0\) for all \(I \in \, ]0,I_0]\),
which implies \(P^\prime(0) > 1\) for all \(I \in \, ]0,I_0]\).
This means the origin is a repeller, so there exists some
(unknown) \(\mu > 0\) where \(]0,\mu]\) is negatively
invariant for \eqref{eqn:r-discrete}, and all solutions in this
interval flow through \(r = \mu\).
If \(I < 0\), we have \(P^\prime(0) < 1\) and the origin is an
asymptotically stable equilibrium for \eqref{eqn:r-discrete}.
